---
title: "Prediction Assignment Writeup"
author: "Ira Lowe"
date: "May 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Assignment

The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Read Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
library (caret)
library(randomForest)

 pm1train <- read.csv("C:/Users/gatom/Coursera/MachineLearn/pml-training.csv")
 pm1test <- read.csv("C:/Users/gatom/Coursera/MachineLearn/pml-testing.csv")
 
```
#Explore Data

``` {r }
dim(pm1train)  

head(pm1train,3)  
```
#Find predictors
```{r}
#Find predictors
#eliminate first few non predictor columns and remove NA and near zero variables

n0var <- nearZeroVar(pm1train)
TrainSet1 <- pm1train[, -n0var]
nmissing  <- sapply(TrainSet1, function(x) mean(is.na(x))) > 0.95
TrainSet2 <- TrainSet1[,nmissing == FALSE] 
TrainSet3 <- TrainSet2[, 7:59]

# Do the Same for Testing
n0var <- nearZeroVar(pm1test)
TestSet1 <- pm1test[, -n0var]
nmissing  <- sapply(TestSet1, function(x) mean(is.na(x))) > 0.95
TestSet2 <- TestSet1[,nmissing == FALSE] 
TestSet3 <- TestSet2[, 7:59]

```
#Run Model Random Forest  and Validation
```{r}
#create validation file

set.seed <-3324
inTrain <- createDataPartition(y=TrainSet3$classe, p=0.7, list = FALSE)
  training <- TrainSet3[inTrain,]
  validation <- TrainSet3[-inTrain,]
#Run Model and Validation

Train.rf <- randomForest(x = training[,-53], y= training[,53], ntrees = 100)
predict.rf <- predict(Train.rf, validation)
confusionMatrix(table(predict.rf, validation$classe))


```
# Predict Exercise classe from Test data
```{r}
predict.test <- predict(Train.rf, TestSet3)

print(predict.test)

```
#Conclusion
The random forest did an excellent job of predicting the exercise and got 20/20 correct.

 
  